Optimizing Abstract Abstract Machines

### 2012-11-14 paper/paper.tex
Finished intro to implementation techniques

We discuss a succession of techniques for increasing performance by
decreasing the interpretive overhead of abstract interpretation. They
come in two flavors: decreasing the size of the state space with domain-specific knowledge [DSK], and
decreasing the amount of intermediate data structures [Eng](ineering).

\begin{enumerate}
 \item{[DSK] lazy non-determinism: quotient by sets of values}
 \item{[DSK] abstract compilation: eliminate {\tt ev} states}
 \item{[Eng] store deltas: operate on store changes rather than whole stores}
 \item{[DSK/Eng] timestamp: represent the store as a global, mutable
   hash and determine a state's store by the number of changes made to
   the global store at the time of visiting}
 \item{[Eng] preallocation: represent the global store as a global, mutable vector}
 \item{[DSK] uniform literal approximation: coarsely abstract literal data}
\end{enumerate}

Some techniques preserve the precision of the underlying analysis, and
others do not. We will discuss the design decisions behind the
precision-lossy techniques.  Each technique requires minimal changes
to the reduction relation, and each improves performance by
integer factors.

>>>>

In this section, we discuss the optimizations for abstract interpreters that
yield our ultimate performance gains.
%
We have two broad categories of these optimizations: (1) transition elimination
and (2) pragmatic improvement.
%
The transition-elimination optimizations reduce the overall number of transitions
made by the analyzer by performing:
\begin{enumerate}
 \item lazy non-determinism;
 \item abstract compilation; and
 \item uniform literal approximation.
\end{enumerate}
The pragmatic improvements reduce overhead and trade space for time
by utilizing:
\begin{enumerate}
 \item store deltas;
 \item timestamped stores; and
 \item preallocated data structures.
\end{enumerate}

Some techniques preserve the precision of the underlying analysis, and others
do not.
%
For any technique that loses precision, we will discuss the design rationale
for the move.
